### 2025-06-19 simulation.qmd
20250619 - Monte Carlo simulation

bootstrappedFantasyPoints_summary %>% 
  filter(position == "PK") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

@@

@@

calculate_fantasyPoints <- function(df, scoring_rules){
  #fantasyPoints <-

  #return(fantasyPoints)
}

monte_carlo_simulation <- function(position_df, num_simulations = 10000){
  player_ids <- unique(position_df$id)

  results <- lapply(player_ids, function(player_id){ # use curly brackets when there are multiple expressions
    player_data <- position_df %>% filter(id == player_id)

    # Run the simulation for this player
    simulations <- replicate(
      num_simulations, {
        sampled_data <- player_data %>%
          summarise(across(
            c(pass_yds, pass_tds, rush_yds, rush_tds, rec_yds, rec_tds), # add more statistical categories
            ~ if (all(is.na(.))) NA else sample(.[!is.na(.)], 1, replace = TRUE)))

        sampled_data # return last expression
        },
      simplify = FALSE)

    simulation_df <- dplyr::bind_rows(simulations)
    #simulation_df$fantasyPoints <- calculate_fantasyPoints(simulation_df) # write function to calculate fantasy points

    ci <- simulation_df %>%
      summarise(across(where(is.numeric), list(
        lower = \(x) quantile(x, 0.025, na.rm = TRUE),
        upper = \(x) quantile(x, 0.975, na.rm = TRUE))))

    ci <- ci %>% 
      mutate(id = player_id)

    return(ci)
    })

    # Combine results for all players
    final_ci <- bind_rows(results)

    return(final_ci)
}

monte_carlo_simulation(players_projectedPoints_seasonal$QB, num_simulations = 10)

QB_ci <- monte_carlo_simulation(players_projectedPoints_seasonal$QB)
```

>>>>

bootstrappedFantasyPoints_summary %>% 
  filter(position == c("K","PK")) %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

@@

#### Prepare Data {#sec-simulationProjectedPointsMonteCarloPrepareData}

```{r}
#| include: false

all_proj <- dplyr::bind_rows(players_projectedPoints_seasonal)
```

```{r}
all_proj <- all_proj %>% 
  rename(projectedPoints = raw_points)
```

#### Optimal Distribution for Each Player {#sec-simulationProjectedPointsMonteCarloOptimalDistribution}

```{r}
# Function to identify the "best" distribution (Normal vs Skew‑Normal) for every player (tries both families and picks by AIC; uses empirical distribution if fewer than 2 unique scores)
fit_best <- function(x) {
  # Basic checks
  if (length(unique(x)) < 2 || all(is.na(x))) { # Use empirical distribution if there are fewer than 2 unique scores
    return(list(type = "empirical", empirical = x))
  }

  # Try Normal Distribution
  fit_norm <- tryCatch(
    fitdistrplus::fitdist(x, distr = "norm"),
    error = function(e) NULL
  )

  # Try Skew-Normal Distribution
  fit_skew <- tryCatch(
    sn::selm(x ~ 1),
    error = function(e) NULL
  )

  # Handle bad fits: sd = NA, etc.
  if (!is.null(fit_norm) && any(is.na(fit_norm$estimate))) {
    fit_norm <- NULL
  }

  if (!is.null(fit_skew)) {
    pars <- tryCatch(sn::coef(fit_skew, param.type = "dp"), error = function(e) NULL)
    if (is.null(pars) || any(is.na(pars))) {
      fit_skew <- NULL
    }
  }

  # Choose best available
  if (!is.null(fit_norm) && !is.null(fit_skew)) {
    aic_norm <- AIC(fit_norm)
    aic_skew <- AIC(fit_skew)
    if (aic_skew + 2 < aic_norm) { # skew-normal is more complex (has more parameters) than normal distribution, so only select a skew-normal distribution if it fits substantially better than a normal distribution
      pars <- sn::coef(fit_skew, param.type = "dp")
      return(list(
        type  = "skewnorm",
        xi    = pars["dp.location"],
        omega = pars["dp.scale"],
        alpha = pars["dp.shape"]))
    } else {
      return(list(
        type = "norm",
        mean = fit_norm$estimate["mean"],
        sd   = fit_norm$estimate["sd"]))
    }
  } else if (!is.null(fit_norm)) {
    return(list(
      type = "norm",
      mean = fit_norm$estimate["mean"],
      sd   = fit_norm$estimate["sd"]))
  } else {
    return(list(type = "empirical", empirical = x))
  }
}
```

```{r}
proj_dists_tbl <- all_proj |>
  filter(!is.na(id) & id != "") |>
  group_by(id) |>
  summarise(
    dist_info = list(fit_best(projectedPoints)),
    n_proj = n(), # record of how many sources they have
    .groups = "drop"
  )

proj_dists <- proj_dists_tbl |>
  filter(!is.na(id) & id != "") |>
  distinct(id, .keep_all = TRUE) |>
  (\(x) setNames(x$dist_info, x$id))()
```

```{r}
proj_dists_tbl |>
  dplyr::mutate(
    dist_type = purrr::map_chr(dist_info, ~ .x$type)
  ) |>
  dplyr::count(dist_type)
```

#### `SimDesign` Step 1: Design Grid {#sec-simulationProjectedPointsMonteCarloDesign}

Now we build the `SimDesign` design grid based on the number of projections that each player had.

```{r}
Design <- proj_dists_tbl |>
  transmute(
    id,
    n_sources = n_proj # how many projections the player had
  )
```

```{r}
#| eval: false

missing_ids <- setdiff(Design$id, names(proj_dists))
length(missing_ids) # should be 0

any(is.na(proj_dists_tbl$id)) # should be FALSE
any(is.na(Design$id))         # should be FALSE
any(is.na(names(proj_dists))) # should be FALSE
```

#### `SimDesign` Step 2: Generate {#sec-simulationProjectedPointsMonteCarloGenerate}

```{r}
Generate <- function(condition, fixed_objects = NULL) {

  dist_info <- fixed_objects$proj_dists[[as.character(condition$id)]]
  n_sources <- condition$n_sources

  sim_points <- switch(
    dist_info$type,
    empirical = sample(
      dist_info$empirical,
      n_sources,
      replace = TRUE),

    norm = rnorm(
      n_sources,
      mean = dist_info$mean,
      sd = dist_info$sd),

    skewnorm = sn::rsn(
      n_sources,
      xi = dist_info$xi,
      omega = dist_info$omega,
      alpha = dist_info$alpha),

    stop("Unknown distribution type: ", dist_info$type)
  )

  data.frame(
    id = condition$id,
    sim_points = sim_points)
}
```

#### `SimDesign` Step 3: Analyze {#sec-simulationProjectedPointsMonteCarloAnalyze}

```{r}
Analyse <- function(condition, dat, fixed_objects = NULL) {
  tibble::tibble(
    id        = condition$id,
    mean_pts  = mean(dat$sim_points, na.rm = TRUE),
    sd_pts    = sd(dat$sim_points, na.rm = TRUE),
    q10       = quantile(dat$sim_points, 0.10, na.rm = TRUE),
    q90       = quantile(dat$sim_points, 0.90, na.rm = TRUE),
    p100      = mean(dat$sim_points >= 100, na.rm = TRUE),
    p150      = mean(dat$sim_points >= 150, na.rm = TRUE),
    p200      = mean(dat$sim_points >= 200, na.rm = TRUE),
    p250      = mean(dat$sim_points >= 250, na.rm = TRUE),
    p300      = mean(dat$sim_points >= 300, na.rm = TRUE),
    p350      = mean(dat$sim_points >= 300, na.rm = TRUE)
  )
}
```

#### `SimDesign` Step 4: Summarize {#sec-simulationProjectedPointsMonteCarloSummarize}

```{r}
Summarise <- function(condition, results, fixed_objects = NULL) {
  dplyr::summarise(
    results,
    across(
      where(is.numeric),
      list(
        mean = ~mean(.x, na.rm = TRUE),
        sd   = ~sd(.x,  na.rm = TRUE)),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )
}
```

#### `SimDesign` Step 5: Run the Simulation {#sec-simulationProjectedPointsMonteCarloRun}

Now, we can run the model.

::: {#nte-monteCarloSimulation .callout-note title="Monte Carlo Simulation"}
Note: the following code that runs the simulation takes a while.
If you just want to save time and load the results object instead of running the simulation, you can load the results object of the simulation (which has already been run) using this code:

```{r}
load(url("https://osf.io/download/ues7n/"))
```
:::

```{r}
#| eval: false

monteCarloSim_results <- SimDesign::runSimulation(
  design = Design,
  replications = 1000,
  generate = Generate,
  analyse = Analyse,
  summarise = Summarise,
  fixed_objects = list(proj_dists = proj_dists),
  seed = genSeeds(Design, iseed = 52242), # for reproducibility
  parallel = TRUE # for faster (parallel) processing
)
```

#### Simulation Results {#sec-simulationProjectedPointsMonteCarloResults}

```{r}
monteCarloSim_results <- monteCarloSim_results %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","position","team")],
    by = c("id" = "mfl_id")
  ) %>% 
  select(name, team, position, everything()) %>% 
  arrange(-mean_pts_mean)
```

@@

save(
  monteCarloSim_results,
  file = "./data/monteCarloSim_results.RData")
```

```{r}
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/monteCarloSim_results.RData", fsep = ""))
```

```{r}
monteCarloSim_results
```

```{r}
monteCarloSim_results %>% 
  filter(position == "QB")
```

```{r}
monteCarloSim_results %>% 
  filter(position == "RB")
```

```{r}
monteCarloSim_results %>% 
  filter(position == "WR")
```

```{r}
monteCarloSim_results %>% 
  filter(position == "TE")
```

```{r}
monteCarloSim_results %>% 
  filter(position %in% c("K","PK"))
```

```{r}
monteCarloSim_results %>% 
  filter(position %in% c("DL","DT","DE"))
```

```{r}
monteCarloSim_results %>% 
  filter(position %in% c("LB","MLB","OLB"))
```

```{r}
monteCarloSim_results %>% 
  filter(position %in% c("DB","S","CB"))
```
